### 2008-12-21 talk/ecoop2009/benchmarks.tex
typo
minor modifications

In section \ref{sec:tlc-properties}, we saw that TLC provides most of the
features that usaully make dynamically typed language so slow, such as
\emph{stack-based interpreter}, \emph{boxed arithmetic} and \emph{dynamic lookup} of
methods and attributes.

In the following sections, we will show some benchmarks that show how our
generated JIT can handle all these features very well.

@@

Moreover, for each benchmark we also show the time taken by running the
equivalent program written in C\# \footnote{The sources for both TLC and C\#
  programs are available at:
  http://codespeak.net/svn/pypy/extradoc/talk/ecoop2009/benchmarks/}

@@

As we can see, the code generated by the JIT can be up to ~1800 times faster
than the non-jitted case.  Moreover, it often runs at the same speed as the
equivalent program written in C\#, being only 1.5 slower in the worst case.

@@

As already discussed, our generated JIT does not compile the whole function at
once. Instead, it compiles and executes code chunk by chunk, waiting until it
knows enough informations to generate highly efficient code.  In particualr,
at the time it emits the code for the inner loop it exactly knows the
type of \lstinline{obj}, thus it can remove the overhead of dynamic dispatch
and inline the method call.  Moreover, since \lstinline{obj} never escapes the
function, it is \emph{virtualized} and its field \lstinline{value} is stored
as a local variable.  As a result, the generated code results in a simple loop
doing additions in-place.

@@

\anto{maybe we should move the following paragraph to
  abstract/introduction/conclusion?}

Despite being only a microbenchmark, this result is very important as it proves
that our strategy of intermixing compile time and runtime can yield to better
performances than current techniques.  The result is even more impressive if
we consider dynamically typed languages as TLC are usually considered much
slower than the statically typed ones.

>>>>

In section \ref{sec:tlc-properties}, we saw that TLC provides most of the
features that usually make dynamically typed language so slow, such as
\emph{stack-based interpreter}, \emph{boxed arithmetic} and \emph{dynamic lookup} of
methods and attributes.

In the following sections, we present some benchmarks that show how our
generated JIT can handle all these features very well.

@@

Moreover, for each benchmark we also show the time taken by running the
equivalent program written in C\#\footnote{The sources for both TLC and C\#
  programs are available at:
  http://codespeak.net/svn/pypy/extradoc/talk/ecoop2009/benchmarks/}

@@

As we can see, the code generated by the JIT can be up to about 1800 times faster
than the non-jitted case.  Moreover, it often runs at the same speed as the
equivalent program written in C\#, being only 1.5 slower in the worst case.

@@

As already discussed, our generated JIT does not compile the whole function at
once. Instead, it compiles and executes code chunk by chunk, waiting until it
knows enough informations to generate highly efficient code.  In particualr,
at the time it emits the code for the inner loop it exactly knows the
type of \lstinline{obj}, thus it can remove the overhead of dynamic dispatch
and inline the method call.  Moreover, since \lstinline{obj} never escapes the
function, it is \emph{virtualized} and its field \lstinline{value} is stored
as a local variable.  As a result, the generated code turns out to be a simple loop
doing additions in-place.

@@

Despite being only a microbenchmark, this result is very important as it proves
that our strategy of intermixing compile time and runtime can yield to better
performances than current techniques.  The result is even more impressive if
we take in account that dynamically typed languages as TLC are usually considered much
slower than the statically typed ones.
